### 2024-03-30 content/chapter_5.tex
:triangular_flag_on_post: Initial submission 240330

\begin{section}{\glsfmtlong{bd}-\glsfmtlong{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ in the following context.
	The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume a lossless asymmetric network\footnote{While symmetric \gls{bd}-\gls{ris} (involving capacitors and inductors) is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2024,Zhou2023,Li2023d,Bartoli2023}, asymmetric \gls{bd}-\gls{ris} can be built over reconfigurable asymmetric passive components (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}.} without mutual coupling between scatter elements, as previously considered in \cite{Li2023b,Li2023c,Bartoli2023}.
	The overall scattering matrix of the \gls{bd}-\gls{ris} is
	\begin{equation}
		\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
	\end{equation}
	where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix that describes the response of group $g \in \mathcal{G} \triangleq \{1, \ldots, G\}$.
	That is, $\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}$.
	Note that diagonal (single-connected) and unitary (fully-connected) \gls{ris} can be regarded as its extreme cases with group size $L=1$ and $L=N_\mathrm{S}$, respectively.
	Some potential physical architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2024}, where the radiation pattern and circuit topology have impacts on the scattering matrix.

	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (transmitter-receiver), backward (\gls{ris}-receiver), and forward (transmitter-\gls{ris}) channels, respectively.
	The equivalent channel is a function of the scattering matrix
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward subchannels for \gls{ris} group $g$, corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
	% That is, $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns (resp. rows) of $\mathbf{H}_\mathrm{B}$ (resp. $\mathbf{H}_\mathrm{F}$).
	% , corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$.
	% $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$
	Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
	\begin{equation}
		\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
	\end{equation}
	where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
	Let $\mathbf{H}_{\mathrm{B},g} = \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \mathbf{V}_{\mathrm{B},g}^\mathsf{H}$ and $\mathbf{H}_{\mathrm{F},g} = \mathbf{U}_{\mathrm{F},g} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}$ be the compact \gls{svd} of the backward and forward channels, respectively.
	The equivalent channel can thus be rewritten as
	\begin{equation}
		\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
		\label{eq:channel_equivalent_svd}
	\end{equation}

	\begin{remark}
		By analyzing \eqref{eq:channel_equivalent_svd}, we conclude that channel shaping through \gls{bd}-\gls{ris} creates two critical opportunities:
		\begin{itemize}
			\item \textbf{Subspace alignment:} each group can align the singular matrices of the backward-forward (intra-group, multiplicative) and the direct-indirect (inter-group, additive) channels. In \gls{siso}, subspace alignment boils down to phase matching and the optimal scattering matrix of group $g$ is
			\begin{equation}
				\mathbf{\Theta}_g^\star = \exp \bigl(\jmath \mathrm{arg}(h_\mathrm{D})\bigr) \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H},
			\end{equation}
			where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr]$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr]$, and $\mathbf{N}_{\mathrm{B},g}, \mathbf{N}_{\mathrm{F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of the null spaces of $\mathbf{h}_{\mathrm{B},g}$ and $\mathbf{h}_{\mathrm{F},g}$, respectively.
			Diagonal \gls{ris} is thus sufficient for perfect \gls{siso} phase matching (with empty null spaces), but its disadvantage in subspace alignment scales with \gls{mimo} dimensions, since each element can only apply a common phase shift to the ``pinhole'' indirect channel of size $N_\mathrm{R} \times N_\mathrm{T}$ passing through itself.
			Even if the \gls{bd}-\gls{ris} is fully-connected, there still exists a trade-off between the direct-indirect and backward-forward alignments.
			% It suggests that diagonal \gls{ris} applying a phase shift to each indirect channel is sufficient for phase matching in \gls{siso}. but its disadvantage in subspace alignment scales with \gls{mimo} dimensions.
			% It suggests that diagonal \gls{ris} applying a phase shift to each indirect channel is sufficient for phase matching in \gls{siso}, but its disadvantage in subspace alignment scales with \gls{mimo} dimensions.
			% It is obvious that diagonal \gls{ris} is sufficient for perfect phase matching. However,
			% It is obvious that diagonal \gls{ris} is sufficient for perfect phase matching. However, in \gls{mimo} each element can only apply a common phase shift to the associated rank-1 $N_\mathrm{R} \times N_\mathrm{T}$ indirect channel.
			% Therefore, perfect subspace alignment of indirect channels through different elements is generally impossible.
			% It means the disadvantage of diagonal \gls{ris} in subspace alignment and subchannel rearrangement scales with \gls{mimo} dimensions.
			% is sufficient for perfect alignment
			% and is sufficient for perfect subspace alignment.
			% In \gls{mimo}, perfect subspace alignment is generally impossible due to the lack of direct control over the indirect channels.aa
			% backward-forward (intra-group, multiplicative) and the direct-indirect (inter-group, additive) alignments of the singular matrices.
			\item \textbf{Subchannel rearrangement:} this unique property of \gls{bd}-\gls{ris} allows each group to rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, diagonal \gls{ris} provides a maximum indirect channel amplitude of $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$, while \gls{bd}-\gls{ris} can generalize it to $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$. By rearrangement inequality, the maximum is attained by pairing the $l$-th strongest backward and forward subchannels in each group.
			The advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions, since the number of subchannels associated with each group grows with $N_\mathrm{R} \times N_\mathrm{T}$.
			% effect is more pronounced in \gls{mimo} since the number of subchannels associated with each group grows quadratically with the dimensions.
			% The advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions. since the number of subchannels associated with each group grows quadratically with the dimensions.
			% It suggests that the advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions.
			% \item \textbf{Subchannel rearrangement:} this unique property of \gls{bd}-\gls{ris} allows each group to rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, the maximum indirect channel amplitude by diagonal \gls{ris} is $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$ while \gls{bd}-\gls{ris} can generalize it to $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are independent permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$. By rearrangement inequality, the maximum is attained by pairing the $l$-th strongest backward and forward subchannels in each group.
			% \begin{equation}
			% 	\item \
			% \end{equation}
			% and $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$ by \gls{bd}-\gls{ris}, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$. By rearrangement inequality,
			% \item \textbf{Subchannel rearrangement:} this unique property of \gls{bd}-\gls{ris} allows each group to rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, the optimal indirect channel amplitude is $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$ by diagonal \gls{ris} and $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},\pi_\mathrm{B}(n)} \rvert \lvert h_{\mathrm{F},\pi_\mathrm{F}(n)} \rvert$ by \gls{bd}-\gls{ris}, where $\pi(\cdot)$ is a permutation of $\{1, \ldots, N_\mathrm{S}\}$.
			% this is unique to \gls{bd}-\gls{ris} --- each group can rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, the optimal indirect channel amplitude is relaxed from $\sum_n \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$ to $\sum_n \lvert h_{\mathrm{B},\pi(n)} \rvert \lvert h_{\mathrm{F},\pi(n)} \rvert$ where $\pi(\cdot)$ is a permutation of $\{1, \ldots, L\}$.
			% from strongest to weakest, which attains the maximum in rearrangement inequality.
		\end{itemize}
	\end{remark}

@@

	% \begin{remark}
	% 	\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size $L$ 1 and $N_\mathrm{S}$, respectively.
	% \end{remark}
	% \begin{remark}
	% 	Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
	% 	This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
	% 	% Later we will show the potential benefits from the \gls{csi} overhead.
	% \end{remark}
\end{section}

\begin{section}{Channel Singular Values Redistribution}
	\begin{subsection}{A Toy Example}\label{sc:toy_example}
		We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
		% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		We also enforce symmetry by $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/chapter_5/singular_trend.eps}
			\caption{$2 \times 2 \times 2$ (no direct) channel singular value shaping by diagonal and symmetry unitary \gls{ris}.}
			\label{fg:singular_trend}
		\end{figure}
		Fig.~\ref{fg:singular_trend} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		It is observed that both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of scattering elements and design parameters.
		A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
		This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values and motivates further studies on channel shaping.
	\end{subsection}

	\begin{subsection}{Pareto Frontier Characterization}
		We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pareto}}{\label{ob:pareto}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{co:pareto_unitary}}
		\end{maxi!}
		where $n \in \{1,\ldots,N \triangleq \min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
		Varying $\{\rho_n\}$ unveils the entire achievable singular value region.
		Thus, the Pareto frontier problem~\eqref{op:pareto} generalizes most relevant metrics and provides a powerful shaping framework.
		The objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}$ and the feasible domain \eqref{co:pareto_unitary} for group $g$ corresponds to the Stiefel manifold.
		Next, we zoom out to general smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.

>>>>

\begin{section}{\glsfmtshort{mimo}-\glsfmtshort{pc}}
	\begin{subsection}{System Model}
		Consider a \gls{bd}-\gls{ris}-aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
		This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ in the following context.
		The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
		To simplify the analysis, we assume a lossless asymmetric network\footnote{While symmetric \gls{bd}-\gls{ris} (involving capacitors and inductors) is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2024,Zhou2023,Li2023d,Bartoli2023}, asymmetric \gls{bd}-\gls{ris} can be built over reconfigurable asymmetric passive components (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}.} without mutual coupling between scatter elements, as previously considered in \cite{Li2023b,Li2023c,Bartoli2023}.
		The overall scattering matrix of the \gls{bd}-\gls{ris} is
		\begin{equation}
			\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
		\end{equation}
		where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix that describes the response of group $g \in \mathcal{G} \triangleq \{1, \ldots, G\}$.
		That is, $\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}$.
		Note that diagonal (single-connected) and unitary (fully-connected) \gls{ris} can be regarded as its extreme cases with group size $L=1$ and $L=N_\mathrm{S}$, respectively.
		Some potential physical architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2024}, where the radiation pattern and circuit topology have impacts on the scattering matrix.

		Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (transmitter-receiver), backward (\gls{ris}-receiver), and forward (transmitter-\gls{ris}) channels, respectively.
		The equivalent channel is a function of the scattering matrix
		\begin{equation}
			\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
			\label{eq:channel_equivalent}
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward subchannels for \gls{ris} group $g$, corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
		% That is, $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns (resp. rows) of $\mathbf{H}_\mathrm{B}$ (resp. $\mathbf{H}_\mathrm{F}$).
		% , corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$.
		% $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$
		Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\begin{equation}
			\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
			\label{eq:unitary_decomposition}
		\end{equation}
		where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
		Let $\mathbf{H}_{\mathrm{B},g} = \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \mathbf{V}_{\mathrm{B},g}^\mathsf{H}$ and $\mathbf{H}_{\mathrm{F},g} = \mathbf{U}_{\mathrm{F},g} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}$ be the compact \gls{svd} of the backward and forward channels, respectively.
		The equivalent channel can thus be rewritten as
		\begin{equation}
			\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
			\label{eq:channel_equivalent_svd}
		\end{equation}

		\begin{remark}\label{rm:channel_shaping}
			By analyzing \eqref{eq:channel_equivalent_svd}, we conclude that channel shaping through \gls{bd}-\gls{ris} creates two critical opportunities:
			\begin{itemize}
				\item \textbf{Subspace alignment:} Each group can align the singular matrices of the backward-forward (intra-group, multiplicative) and the direct-indirect (inter-group, additive) channels. In \gls{siso}, subspace alignment boils down to phase matching and the optimal scattering matrix of group $g$ is
				\begin{equation}
					\mathbf{\Theta}_g^\star = \exp \bigl(\jmath \mathrm{arg}(h_\mathrm{D})\bigr) \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H},
				\end{equation}
				where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr]$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr]$, and $\mathbf{N}_{\mathrm{B},g}, \mathbf{N}_{\mathrm{F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of the null spaces of $\mathbf{h}_{\mathrm{B},g}$ and $\mathbf{h}_{\mathrm{F},g}$, respectively.
				Diagonal \gls{ris} is thus sufficient for perfect \gls{siso} phase matching (with empty null spaces), but its disadvantage in subspace alignment scales with \gls{mimo} dimensions, since each element can only apply a common phase shift to the ``pinhole'' indirect channel of size $N_\mathrm{R} \times N_\mathrm{T}$ passing through itself.
				Even if the \gls{bd}-\gls{ris} is fully-connected, there still exists a trade-off between the direct-indirect and backward-forward alignments.
				\item \textbf{Subchannel rearrangement:} This unique property of \gls{bd}-\gls{ris} allows each group to rearrange and combine the backward and forward subchannels by strength. In \gls{siso}, diagonal \gls{ris} provides a maximum indirect channel amplitude of $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$, while \gls{bd}-\gls{ris} can generalize it to $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$. By rearrangement inequality, the maximum is attained by pairing the $l$-th strongest backward and forward subchannels in each group.
				The advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions, since the number of subchannels associated with each group grows with $N_\mathrm{R} \times N_\mathrm{T}$.
			\end{itemize}
		\end{remark}

@@

		% \begin{remark}
		% 	\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size $L$ 1 and $N_\mathrm{S}$, respectively.
		% \end{remark}
		% \begin{remark}
		% 	Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
		% 	This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
		% 	% Later we will show the potential benefits from the \gls{csi} overhead.
		% \end{remark}
	\end{subsection}

	\begin{subsection}{Channel Singular Values Redistribution}
		\begin{subsubsection}{A Toy Example}\label{sc:toy_example}
			We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
			% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
			Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
			The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
			\begin{equation}
				\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
					e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
					-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
				\end{bmatrix}.
			\end{equation}
			It is worth noting that $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
			For a fair comparison, we also enforce symmetry $\mathbf{\Theta}_\mathrm{U} = \mathbf{\Theta}_\mathrm{U}^\mathsf{T}$ by $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
			\begin{figure}
				\centering
				\includegraphics[width=\columnwidth]{assets/chapter_5/singular_trend.eps}
				\caption{$2 \times 2 \times 2$ (no direct) channel singular value shaping by diagonal and symmetry unitary \gls{ris}.}
				\label{fg:singular_trend}
			\end{figure}
			Fig.~\ref{fg:singular_trend} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
			It is observed that both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of scattering elements and design parameters.
			A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
			This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values and motivates further studies on channel shaping.
		\end{subsubsection}

		\begin{subsubsection}{Pareto Frontier Characterization}\label{sc:pareto_frontier}
			We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
			\begin{maxi!}
				{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pareto}}{\label{ob:pareto}}
				\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I},}{\quad \forall g,}{\label{co:pareto_unitary}}
			\end{maxi!}
			where $n \in \{1,\ldots,N \triangleq \min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
			Varying $\{\rho_n\}$ unveils the entire achievable singular value region.
			Thus, we claim the Pareto frontier problem~\eqref{op:pareto} generalizes most relevant metrics and provides a powerful shaping framework.
			The objective \eqref{ob:pareto} is smooth in $\mathbf{\Theta}$ and the feasible domain \eqref{co:pareto_unitary} for group $g$ corresponds to the Stiefel manifold.
			Next, we zoom out and provide an efficient optimization framework for asymmetric \gls{bd}-\gls{ris}.

			Inspired by \cite{Abrudan2008,Abrudan2009}, we propose a block-wise \gls{rcg} algorithm along the geodesics on the Lie group of unitary matrices $\mathbb{U}^{L \times L}$.
			It leverages the fact that unitary matrices are closed under multiplication.
			At iteration $r$, the gradient is computed in the Euclidean space and translated to the Riemannian manifold \cite{Absil2009}
			\begin{gather}
				\nabla_{\mathrm{E},g}^{(r)} = \frac{\partial f(\mathbf{\Theta}_g^{(r)})}{\partial \mathbf{\Theta}_g^*},\label{eq:gradient_euclidean}\\
				\nabla_{\mathrm{R},g}^{(r)} = \nabla_{\mathrm{E},g}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)}}^\mathsf{H}.\label{eq:gradient_riemannian} % $\mathcal{O}(N_\mathrm{S}^3)$
			\end{gather}
			The Polak-Ribierre parameter \cite{Polak1969} is approximated as \cite{Abrudan2009}
			\begin{equation}
				\gamma_g^{(r)} = \frac{\mathrm{tr}\bigl((\nabla_{\mathrm{R},g}^{(r)} - \nabla_{\mathrm{R},g}^{(r-1)}) {\nabla_{\mathrm{R},g}^{(r)}}^\mathsf{H}\bigr)}{\mathrm{tr}\bigl(\nabla_{\mathrm{R},g}^{(r-1)} {\nabla_{\mathrm{R},g}^{(r-1)}}^\mathsf{H}\bigr)}, % $\mathcal{O}(2 N_\mathrm{S}^3 + N_\mathrm{S}^2 + 2 N_\mathrm{S})$
				\label{eq:parameter_cg}
			\end{equation}
			and the conjugate direction is
			\begin{equation}
				\mathbf{D}_g^{(r)} = \nabla_{\mathrm{R},g}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)}. % $\mathcal{O}(N_\mathrm{S}^2)$
				\label{eq:direction_cg}
			\end{equation}
			In the Stiefel manifold, the geodesic emanating from $\mathbf{\Theta}_g^{(r)}$ with velocity $\mathbf{D}_g^{(r)}$ and step size $\mu$ is described compactly by the exponential map \cite{Edelman1998}
			\begin{equation}
				\mathbf{G}_g^{(r)}(\mu) = \exp(\mu \mathbf{D}_g^{(r)}) \mathbf{\Theta}_g^{(r)}. % $\mathcal{O}(N_\mathrm{S}^3)$
				\label{eq:geodesic}
			\end{equation}
			An appropriate $\mu^\star$ can be obtained by the Armijo rule \cite{Armijo1966}.
			To double the step size, one can simply square the rotation matrix instead of recomputing the matrix exponential, i.e., $\exp(2 \mu \mathbf{D}_g^{(r)}) = \exp^2(\mu \mathbf{D}_g^{(r)})$.
			To halve the step size, taking its square root is sufficient in theory but is subject to numerical instability in practice.
			Finally, the scattering matrix is updated along the geodesic as
			\begin{equation}
				\mathbf{\Theta}_g^{(r+1)} = \mathbf{G}_g^{(r)}(\mu^\star).
				\label{eq:update_geodesic}
			\end{equation}
			Algorithm~\ref{ag:rcg} summarizes the proposed block-wise geodesic \gls{rcg} method for smooth maximization problems of asymmetric \gls{bd}-\gls{ris}.
			Convergence to stationary points is guaranteed.
